\chapter{Longitudinal Glaucomatous Visual Field Dataset} \label{chapter:rotterdam}

The Longitudinal Glaucomatous Visual Field dataset is publicly available \href{http://www.rodrep.com/}{online from Rotterdam Ophthalmic Institute}. \cite{Bryan2013}

The following is an excerpt from the Rotterdam dataset information:

\begin{quotation}
The data set contains longitudinal visual field data of 139 glaucoma patients. These patients were recruited from the Rotterdam Eye Hospital (Rotterdam, The Netherlands). Informed consent was obtained from all subjects.

Inclusion criteria:

\begin{itemize}
\item Between 18 and 85 years at time of inclusion.
\item Glaucoma diagnosis: Two of the following conditions are met: pattern standard deviation significant at p=0.05, abnormal hemifield test result, cluster of $\leq3$ points depressed at p=0.05 level or 1 point at p=0.01. VF defects must be reproducible on at least one occasion.
\item Primary open-angle glaucoma, normal tension glaucoma, pigmentary glaucoma, and treated angle closure glaucoma. (Pseudo-exfoliation glaucoma was not included.)
\end{itemize}


Exclusion criteria:

\begin{itemize}
\item Secondary glaucomas except pigmentary.
\item Evidence of SAP VF abnormality consistent with other disease.
\item BCVA $> 0.3$ (LogMAR).
\item Refractive error outside -10.0 to +5.0 D range.
\item Cataract surgery in previous 12 months.
\item Previous refractive or vitreoretinal surgery.
\item Evidence of diabetic retinopathy, diabetic macular edema, or other vitreo-retinal disease.
\item Previous keratoplastic surgery.
\item Diabetes, leukemia, AIDS, uncontrolled systemic hypertension, multiple sclerosis or (other) life threatening disease.
\end{itemize}


Both eyes of each participant were included. Visits were scheduled every 6 months. At each visit, standard clinical ophthalmic examinations were performed, including visual acuity, intra-ocular pressure, gonioscopy and ophthalmoscopy. At each visit, standard automated perimetry was also performed. Visual fields were acquired on a Humphrey Visual Field Analyzer (Carl Zeiss Meditec) with a standard white-on-white 24-2 field with the full threshold program. The provided data set contains information on the visual field and on the individual visual field test locations.

Included data:

\begin{itemize}
\item Patient's gender
\item Visit data (per eye): Patient's age at visit, IOP, MD.
\item Local visual field data (54 points per field): Threshold, total deviation
\end{itemize}

\end{quotation}

